A decentralized network support middleware for MMOGs

### 2009-09-09 textos/PlanoDoutorado/Lugano/short_planodoc_lugano.tex

Massively multiplayer online games (MMOGs) are the theme of this study plan. Usually a client-server infrastructure is used to support this kind of games. This plan proposes the investigation and definition of a novel distribution model to be used for MMOGs, reducing the maintenance cost of the game server. In the next sections, the state of the art regarding this area will be presented.

\section{Multiplayer games}

% o que são
% breve histórico e exemplos
% requisitos computacionais (sem mtos detalhes)

In the past few decades, since the creation of the personal computer, computer games have enjoyed an evergrowing popularity. After the acess to the Internet became widely available to the general public, with increasingly fast and cheap domestic connections, multiplayer online games have become an important market. To illustrate this, a study from 2006 estimated that the online games market would grow from US\$ 3.4 billion in 2005 to US\$ 13 billion in 2011 \cite{intelligence2006online}. The main characteristic of the MMOGs is the large number of concurrent players interacting with one another through the Internet. As examples of well-known online multiplayer games, we may cite \emph{Quake} \cite{quake} as a classic FPS game; \emph{Starcraft: Brood War} \cite{broodwar} a classic RTS game still played nowadays and Diablo II \cite{diablo2} as an RPG example.

\section{Massively multiplayer online games (MMOGs)}
%TODO: por um monte de referências a MMOGs, do tipo (de maior sucesso são X, Y, Z), outros exemplos são A, B, C
% requisitos computacionais elevados a 1000

After single-player computer games evolved to multiplayer online games with the advent of the Internet, MMOGs emerged as a new trend because of the decrease of the domestic Internet connection cost and the increase of its average bandwidth. This kind of games have become very popular in the latest years because they allow the online interaction of a massive number of players at the same time. In the most succesful cases, such as World of Warcraft \cite{worldofwarcraft}, Lineage II \cite{lineage2} and EVE Online \cite{eveonline}, for example, tens of thousands of players are supported simultaneously \cite{chen2006pch}. These players can, then, interact with one another and with the virtual environment of the game.

One of the main features of these games, present in almost all commercial MMOGs, is that the matches are very long. In some MMORTS (real-time strategy MMOGs), such as Travian \cite{travian}, the games come to last from six months to one year. MMORPGs (MMOGs in RPG style), as is the case of World of Warcraft, the games have no given time to end. The worlds of these games are so vast and populous -- as regards the number of players -- that they may evolve and modify their state independently of some players being playing it or not, because there is always someone connected so the game never stops. For this reason, it is expected that each player can disconnect and continue playing later, without having to start from scratch. Moreover, it is expected that changes made on the game world persist, no matter how many times each player enters and leaves.

Generally, each player controls an entity called \emph{avatar}, which is nothing more than his representation in the virtual environment -- a character who executes the orders given by the player, therefore interfering in the history and in the outcome of the game. For the game to remain consistent for the different participants, each action performed by each avatar must be processed by the server, which calculates the game state resulting from such action. This new state is then broadcast to other players. Other changes in the game world, such as destruction or creation of objects in the environment, and events such as actions of characters controlled by the server must also be sent to the players involved. It is then through these messages sent by the player to the server and in the opposite direction that players can interact with each other and with the simulated world of the game.

However, it is perceived that such a mechanism can easily saturate the bandwidth and overload the cpu(s) of the server. What is usually done is build an infrastructure with a large central server, usually a cluster with its nodes connected through a local high-speed and low latency network, with an Internet connection of some hundreds or thousands of MBps. The greatest cost comes from maintaining this connection, whose main purpose is to maintain every player as up-to-date as possible about the game state -- naturally, tolerating some delay, which should not be over a given threshold.

%Jogos dessa natureza têm alguns requisitos \cite{schiele2007rpp}, dentre os quais podem ser citados: \textbf{persistência} do estado do mundo virtual e dos avatares; \textbf{disponibilidade} do serviço; \textbf{consistência} do estado do mundo e dos avatares entre os diferentes participantes; \textbf{resistência a trapaça}, no que se refere a tentativas de subversão das regras do jogo por jogadores desonestos e \textbf{escalabilidade}, tendo em vista que o grande número de jogadores poderá facilmente sobrecarregar a infra-estrutura de suporte se esta não tiver capacidade de processamento e, principalmente, largura de banda suficientes.

\section{Studying in Switzerland}

@@

>>>>

\chapter*{Study plan summary}

This study plan proposes a research project focused on the theme of massively multiplayer online games (MMOGs). Although this kind of games is usually deployed on a client-server infrastructure, with a large and expensive central server, it is possible to use a network with a decentralized architecture. Therefore, ways of providing such decentralized support should be investigated, reducing the maintenance cost of these games, while still providing the basic requirements of MMOGs.

\section*{Context}

Some years ago, with the broader access to the Internet, the traditional computer games have evolved to multiplayer online games, but with no much more than ten or twenty players in a single match. After that, in the past few years, with faster, cheaper -- when compared to a decade ago --, and more reliable Internet connections, a new game genre has arised as a new trend, which is the MMOG, with potentially tens of thousans of simultaneous players. Some studies demonstrate that the market share of these massively multiplayer games will keep growing at a considerable rate -- according to \cite{intelligence2006online}, the online games market will grow from US\$ 3.4 billion in 2005 to US\$ 13 billion in 2011.

The main reason for the success of these games is the kind of interaction they allow. Usually, such games contain a virtual environment, where its numerous players may interact in many different ways, from fighting -- cooperatively, against artificial intelligence controlled monsters, or even dueling, and fighting each other -- to trading in-game objects -- like weapons and armor in medieval fantasy games -- or just chatting in some place of the game virtual world. This game world, besides, is persistent, i.e. the changes resulting from the player actions will persist, no matter if they keep playing continuously or if they disconnect from the game for a while.

This persitence constitutes one of the main features of these games, whose matches usually do not have any predetermined time to end. There are some exceptions, such as MMORTS (real-time strategy MMOGs), with matches lasting from six months to one year \cite{travian}, what is still considerably long for a game. Most MMOGs, however, have no such time to finish, creating the illusion of a parallel world into which the players may transport themselves when they so desire.

Each player interacts with the virtual environment -- or game world -- by controlling an entity called \emph{avatar}. Each avatar is a representation of a player in the virtual environment -- e.g. a sniper in a war game, a medieval knight in a fantasy game, a gangster or any other character controlled by the player. It performs the actions requested, interfering in the game world and potentially change its state and the state of other avatars. In the traditional central server model, each player sends his desired actions to the server, which processes them, calculating their outcome, represented as new states for the world and for some avatars. This new state is then broadcast to other players, so they can perceive the state changes. The server also acts as an arbiter, who decides the actions performed by non-playable characters, whose state is also sent to the players.

It is easy to observe that the central server, receiving actions from thousands of players and broadcasting the resulting state of each of these actions to, potentially, every player, is prone to have its processing power and, specially, its bandwidth saturated by the game. The machine serving the game is, usually, a cluster consisting of nodes connected via a fast and low-latency local network, providing the necessary processing however, with a very fast Internet, ranging from hundreds of MBps to several GBps. Obviously, such infrastructure is highly expensive, restricting the MMOG market to small game developing companies or research groups with low financial resources.

\section*{Studying in Switzerland}

@@

\section*{Work schedule}
